### 2012-06-20 paper/chemokine.tex
Fixed up the ABM discussion.  ABM allows us to visuall observe spatial effects.

\begin{figure}[ht!]
\begin{center}
 \includegraphics[width=0.7\textwidth]{tcellratio}
 \end{center}
\caption{A logarithmic plot of the ratio of T-Cells at the focus of infection vs the number of cells that make up the focus of infection.  aH5N1 exhibits unstable behavior once the infection is and T-cell counts are reduced to negligible levels.} 
 \label{fig:tcellratio}
\end{figure}

@@

@@

\subsection*{Model Validation}

The ABM produces infection dynamics consistent with \textit{in vivo} observations.  We observe FOI diameters on the order of 1mm in seasonal H1N1 simulation.  The T-cell response at day five post-infection rapidly decreases the number of infected cells over the first 24 hours.

Scaling to humans?

The use of an ABM has advantages over a spatially homogeneous ODE model.  An ODE model assumes that any virus particle is capable of infecting any healthy cell.  Figure \ref{fig:cycells} shows that this is clearly not the case.  In fact, most free virus exists on top of infected cells that are no longer candidates for viral binding and fusion.  ODE models account for this discrepancy by lowering rates of infection, but this simply results in improper parameter values.

The spatial nature of the simulation also reveals challenges to T-cell detection rates that are not apparent in ODE models.  T-cells find infected cells by climbing the chemokine gradient.  This results in T-cell clustering at local maxima of chemokine concentration.  Thus, T-cells do not increase in effectiveness as the number of T-cells increases.  Furthermore, infected cells grow more disperse as the size of the infection grows.  Because T-cells cluster together they become incapable of covering the increasing plaque size.  This allows for the infection to recover and accelerate once its size crosses a certain threshold.  This is another effect that is not seen in an ODE model.

>>>>

\begin{figure}[ht!]
\begin{center}
 \includegraphics[width=0.7\textwidth]{tcellratio}
 \end{center}
\caption{A logarithmic plot of the ratio of T-Cells at the focus of infection vs the number of cells that make up the focus of infection.  aH5N1 exhibits unstable behavior once the infection is cleared and T-cell counts are reduced to negligible levels.} 
 \label{fig:tcellratio}
\end{figure}

@@

The use of an ABM has advantages over a spatially homogeneous ODE model.  An ODE model assumes that any virus particle is capable of infecting any healthy cell.  Figure \ref{fig:cycells} shows that this is clearly not the case.  In fact, most free virus exists on top of infected cells that are no longer candidates for viral binding and fusion.  ODE models account for this discrepancy by lowering rates of infection by a constant amount, but this assumes that the proportion of unsuitable virions will always be the same.  This is clearly false as can been seen in Figure~\ref{fig:cycells} where the early infection has a higher proportion of virus overlapping healthy cells.

Another advantage of our ABM is the ability to render the model using OpenGL (Fig.~\ref{fig:cycells}).  The ability to see the model in action reveals spatial dynamics that are absent in ODE models and difficult or impossible to observe in \textit{in vitro} and \textit{in vivo} systems.  For example, observing the simulation reveals challenges to T-cell detection of infected cells that are not apparent using standard population plots.  By watching the model progress, we observed three key phenomena.  First, because T-cells find infected cells by climbing the chemokine gradient, we see T-cell clustering at local maxima of chemokine concentration.  Thus, we have an explanation for why T-cells do not increase in effectiveness as their numbers increase.  Second, we notice that clumps of T-cells remain in place even after all virus-secreting cells have been killed off.  This is because it takes time for the local chemokine maxima to diffuse and decay to a point where the T-cells can once again climb the gradient to a new maxima.  Finall, we can see that infected cells grow more disperse as the size of the infection grows.  Because T-cells cluster together they become incapable of covering the increasing plaque size.  These spatial observations provide explanations for the pH1N1 resurgence that would be hidden without the visualization tools provided by the ABM.

\subsection*{Model Comparison}

@@
